\textbf{Adaptation of the PHQ-4 Short Screening for Depression and Anxiety to increase its Sensitivity to Subclinical Variability}

### 2023-01-17 paper/2_JAD/manuscript.tex
update JAD

\usepackage{xpatch}
\makeatletter
\xapptocmd\appendix
  {\xapptocmd\section
    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
    {}{\InnerPatchFailed}%
  }
{}{\PatchFailed}
\keywords{PHQ-4, depression, anxiety, brief questionnaire, short scale\newline\indent Word count: 1158}
\usepackage{lineno}

\linenumbers
\usepackage{csquotes}
\usepackage[titles]{tocloft}
\cftpagenumbersoff{figure}
\renewcommand{\cftfigpresnum}{\itshape\figurename\enspace}
\renewcommand{\cftfigaftersnum}{.\space}
\setlength{\cftfigindent}{0pt}
\setlength{\cftafterloftitleskip}{0pt}
\settowidth{\cftfignumwidth}{Figure 10.\qquad}
\usepackage[labelfont=bf, font={scriptsize, color=gray}]{caption}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Adaptation of the PHQ-4 Short Screening for Depression and Anxiety to increase its Sensitivity to Subclinical Variability},
  pdfauthor={Dominique Makowski1, An Shu Te1, \& S.H. Annabel Chen1, 2, 3, 4},
  pdflang={en-EN},
  pdfkeywords={PHQ-4, depression, anxiety, brief questionnaire, short scale},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

@@

\abstract{%
The PHQ-4 is an ultra brief (4 items) screening questionnaire for depression and anxiety. We propose to add one aditional response option (``Once or twice'', in between ``Not at all'' and ``Several days'') to improve its sensitivity to milder alterations, and thus increase its usefulness in subclinical populations. Using Item Response Theory (IRT), we show that this new option does indeed capture specific portions of the measured constructs, and provide a scoring procedure that preserve the compatability with past versions.

@@

While the scale has been validated and used in the general population (Hajek \& König, 2020; Löwe et al., 2010), its initial purpose was to reliably discriminate and identify potential MDD/GAD patients. This discriminative goal materializes in the scale's design and the existence of categorical cut-offs, which does not necessary entail a strong sensitivity to milder mood alterations. In particular, the gap between the two lowest possible answers, ``Not at all'' and ``Several days'', is quite large and leaves out the possibility of more subtle occurrences. While this is not necessarily an issue in clinical and diagnostic contexts, it might lead to a sub-optimal discrimination of affective levels on the lower end of the spectrum, important for instance in the context of subclinical variability quantifications. The goal of this study is hence to enhance, with minimal changes to the original scale, the sensitivity to mild mood level inflections.

@@

The sample consists of 485 English-speaking participants (Mean age = 30.1, SD = 10.1, range: {[}18, 73{]}; Sex: 50.3\% females, 49.7\% males) from the general population who were administered the refined PHQ-4 online as part of another study, which data is available in open-access at \url{https://github.com/RealityBending/IllusionGameReliability}. This study was approved by the NTU Institutional Review Board (NTU IRB-2022-187). All participants provided their informed consent prior to participation and were incentivized after completing the study.

@@

\begin{figure}
\includegraphics[width=1\linewidth]{../figures/figure1} \caption{A. Proportion of answers of each type. B. Item Information Curves from IRT showing the coverage by each item and response of the latent dimension. Typically, an optimally informative item would display a large coverage over theta, with each response presenting a narrow coverage (high discrimination between different levels).}\label{fig:unnamed-chunk-2}
\end{figure}

@@

The objective of this study was to test the introduction of a ``Once or twice'' response option to the PHQ-4 to enhance its sensitivity to milder mood fluctuations. The fact that this new response option was the most prevalent response made by participants in our study is in itself evidence for its usefulness. IRT analysis further revealed that this response captures with precision a unique portion of anxiety and depression. Taken together, our results suggest that adding this option response increases the scale's potential to discriminate average mood levels (which are superior to zero) from lower-end extremes (the true zero).

The hypothesized added value in terms of finer sensitivity should be confirmed by cross-validating the PHQ-4R with more comprehensive measures of anxiety and depression. Additionally, evidence comparing the refined version against the original PHQ-4 would be useful to better understand whether the new ``Once or twice'' option pulls participants that would have otherwise answer ``Not at all'' or ``Several days''. Finally, future studies should further corroborate our statistical validation and test the usefulness of this revision on more targeted samples (e.g., clinical populations) to see if it affects the cut-offs proposed in the literature (and whether it allows for a better discrimination of mood disorders), and refine if need be the normalized scoring estimates in larger sample.

@@

>>>>

\usepackage{xpatch}
\makeatletter
\xapptocmd\appendix
  {\xapptocmd\section
    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
    {}{\InnerPatchFailed}%
  }
{}{\PatchFailed}
\keywords{PHQ-4; depression; anxiety; brief questionnaire; short scale\newline\indent Word count: 1249}
\usepackage{lineno}

\linenumbers
\usepackage{csquotes}
\usepackage[titles]{tocloft}
\cftpagenumbersoff{figure}
\renewcommand{\cftfigpresnum}{\itshape\figurename\enspace}
\renewcommand{\cftfigaftersnum}{.\space}
\setlength{\cftfigindent}{0pt}
\setlength{\cftafterloftitleskip}{0pt}
\settowidth{\cftfignumwidth}{Figure 10.\qquad}
\usepackage[labelfont=bf, font={scriptsize, color=gray}]{caption}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Adaptation of the PHQ-4 Short Screening for Depression and Anxiety to increase its Sensitivity to Subclinical Variability},
  pdfauthor={Dominique Makowski1, An Shu Te1, \& S.H. Annabel Chen1, 2, 3, 4},
  pdflang={en-EN},
  pdfkeywords={PHQ-4; depression; anxiety; brief questionnaire; short scale},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

@@

\abstract{%
The PHQ-4 is an ultra-brief (4 items) screening questionnaire for depression and anxiety. We propose to add one additional response option (``Once or twice'', in between ``Not at all'' and ``Several days'') to improve its sensitivity to milder alterations, and thus increase its usefulness in subclinical populations. Using Item Response Theory (IRT), we provide evidence that this new option does indeed capture specific portions of the measured constructs in the general population.

@@

While the scale has been validated and used in the general population and non-clinical samples (Hajek \& König, 2020; Löwe et al., 2010), its initial purpose was to reliably discriminate and identify potential MDD/GAD patients. This discriminative goal materializes in the scale's design and the existence of categorical cut-offs, which does not necessary entail a strong sensitivity to milder mood alterations. In particular, the gap between the two lowest possible answers, ``Not at all'' and ``Several days'', is quite large and leaves out the possibility of more subtle occurrences. While this is not necessarily an issue in clinical and diagnostic contexts, it might lead to a sub-optimal discrimination of affective levels on the lower end of the spectrum, important for instance in the context of subclinical variability quantifications. The goal of this study is hence to enhance, with minimal changes to the original scale, the sensitivity to mild mood level inflections.

@@

The sample consists of 485 English-speaking participants (Mean age = 30.1, SD = 10.1, range: {[}18, 73{]}; Sex: 50.3\% females, 49.7\% males) from the general population recruited via \emph{Prolific}, a crowd-sourcing platform recognized for providing high quality data (Peer et al., 2022). The only inclusion criterion was a fluent proficiency in English to ensure that the task instructions would be well-understood. Note that the participants were administered the refined PHQ-4 online as part of another study, which data is available in open-access at \url{https://github.com/RealityBending/IllusionGameReliability}. This study was approved by the
NTU Institutional Review Board (NTU IRB-2022-187).
All participants provided their informed consent prior to participation and were incentivized after completing the study.

@@

\begin{figure}
\includegraphics[width=1\linewidth]{../../figures/figure1} \caption{A) Proportion of answers of each type. B) Prevalence of answer pairs. C) Item Information Curves from IRT showing the coverage by each item and response of the latent dimension. Typically, an optimally informative item would display a large coverage over theta, with each response presenting a narrow coverage (high discrimination between different levels).}\label{fig:unnamed-chunk-2}
\end{figure}

@@

The objective of this study was to test the introduction of a ``Once or twice'' response option to the PHQ-4 to enhance its sensitivity to milder mood fluctuations. The fact that this new response option was the most prevalent response made by participants in our study is in itself evidence for its usefulness. IRT analysis further revealed that this response captures with precision a unique portion of the variability in the latent factors measured by the instrument. Taken together, our results suggest that adding this option response increases the scale's potential to discriminate average mood levels (which are superior to zero) from lower-end extremes (the true zero).

It is important to emphasize that our study is a statistical validation in the general population, which analysed how different responses relate to one another. It is a first step providing supporting evidence for the proposed change. However, the structural gains that we observed in terms off finer sensitivity should be confirmed 1) by cross-validating the PHQ-4R with more comprehensive measures of anxiety and depression, and 2) testing the usefulness of this revision on more targeted samples (e.g., clinically characterized populations) to investigate the impact on existing cut-offs and on the detection of mood disorders. Additionally, evidence comparing the refined version against the original PHQ-4 would be useful to better understand whether the new ``Once or twice'' option pulls participants that would have otherwise answer ``Not at all'' or ``Several days''.

@@

\leavevmode\vadjust pre{\hypertarget{ref-peer2022data}{}}%
Peer, E., Rothschild, D., Gordon, A., Evernden, Z., \& Damer, E. (2022). Data quality of platforms and panels for online behavioral research. \emph{Behavior Research Methods}, \emph{54}(4), 1643--1662.
